Optical characterization of gaps in directly bonded Si compound optics using infrared spectroscopy

### 2015-02-16 AO_bondedSi.tex
minor edits to first 17 pages of the text

\begin{abstract}
Silicon direct bonding offers flexibility in the design and development of Si optics by allowing manufacturers to combine subcomponents with a potentially lossless and mechanically stable interface. The bonding process presents challenges in meeting the requirements for optical performance because air gaps at the Si interface cause large Fresnel reflections. Even small (35 nm) gaps reduce transmission through a direct bonded Si compound optic by 4\% at $\lambda = 1.25 \; \mu$m at normal incidence. Understanding the optical effects of such gaps and evaluation of various methods for preventing or eliminating them demands that we have a method for measuring not only the two-dimensional extent of the gaps but also their axial extent.  Existing methods for studying gaps provide only poor measures of this axial extent.  We describe a bond inspection method that makes use of precision transmission spectroscopy.  Using spectroscopy, we can detect and measure gaps as small as 8 nm by measuring transmission as a function of wavelength. Our approach involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes. We describe the measurement and analysis techniques and demonstrate the validity of the approach by measuring bond gaps of known depths produced by microlithography.
\end{abstract}

@@

\subsection{The Si$-$Si interface gap is a low Finesse Fabry-P\`{e}rot etalon}

@@

which has an average value of about 53\%.  Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.  Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqnAbsDSPtrans} in the wavelength range $1200-2500\;$nm.  If the Si wafer path thickness is less than the coherence length for the given spectral bandwidth, then our assumptions break down and the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.  We do not treat this case here since our focus is on thick substrates with small gaps between the interfaces.

\begin{figure}[htbp]
\includegraphics[width=0.40\columnwidth]{figs/etalon_trans.pdf}
\caption{\label{figAbsoluteTrans}The family of normal-incidence curves for transmission as a function of wavelength for various gap sizes in bonded Si optics.  \emph{Top panel}: Absolute transmission through a Si substrate with no gap (top denim-blue solid curve), and small gaps of axial extent 20 nm (green dashed curve), 35 nm (red dash-dotted curve), and 60 nm (purple dotted curve).  The absolute transmission through Si depends on wavelength since the refractive index depends minutely on wavelength.  The curves assume that multiple reflections within the Si substrates add incoherently as discussed in detail in Appendix \ref{sec:Append-IMRTMM}.  \emph{Middle panel}: The ``Gap Transmission'', defined as the Absolute transmission through a sample normalized by the absolute transmission through a gapless sample. The gap sizes and colors are the same as in the top panel.  \emph{Bottom panel}: ``Gap Transmission'' for larger gaps, spanning 0-3500 nm.  The 3500 nm sized gap (cyan solid line), demonstrates characteristic Fabry-P\`erot etalon fringes.}
\end{figure}

@@

The Cary 5000 has four operation modes: single front, single back, double, and double reverse.  We experimented with single front and double modes.  Double mode takes a reference spectrum simultaneously with the target spectrum.  In both single and double mode, a baseline scan is taken with no sample in the measurement holder.  We automatically divide sample spectra by the baseline spectrum.  In single beam mode, the baseline measurements must be repeated about every half hour for post-process division.  In double beam mode the front beam is automatically divided by the rear beam, which corrects out lamp luminosity drift, for example.

@@

The spectral sampling was ranged from 2.0 to 10.0 nm.  We used a spectral resolution of 5.0 nm.  We expect smoothly varying spectral features for all conceivable interface gaps in Si bonded wafers.  Specifically, Fabry-P\`erot fringes attributable to interfacial gaps in bonded Si will be well-sampled with 5 nm spectral resolution, so long as the interface gap is less than $\sim$100 $\mu$m.  The axial extents of such large gaps would be easily measurable by other means, like IR imaging.

@@

We also used IR imaging to inspect VG09-12 and VG15-13.  We detected IR bubbles in all 3 of these bonded Si samples.  The observable bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.  The individual IR bubble diameters were as small as a few square millimeters up to $400\;mm^2$.  The IR bubbles had varied morphologies.  The structure within the IR bubbles was sometimes indicative of Newton's rings, exposing the presence of gaps larger than about $\lambda/2$ up to $\sim 15 \lambda$.  As expected, the unbonded DSP wafers show no such IR bubbles.  No attempt at quantitative measurement was made on the IR images due to evidence of a non-linear detector response.  The bulk morphologies were clear regardless of the detector response.  When IR imaging was available, it was leveraged to target IR specectroscopy in locations demonstrating presence or absense of IR bubbles, as desired.  In other cases when IR imaging was not available or ignored, spectra were taken at random positions on the wafer.

@@

where $f$ is the areal fill factor of the region exhibiting a finite gap size, and $1-f$ is therefore the areal fill factor of the perfectly bonded area.  $T_e$ is defined in equation \ref{eqn:Tetalon}.  The only free parameters in Equation \ref{eqnMix} are $d$ and $f$.  Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tab_meshPatterns}.  These synthetic imbedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secKnownGaps}).

Equipped with an observed spectrum and our mixture model, we constructed the Likelihood function \cite{2013sdmm.book.....I}.  The measurement errors are not assumed to be independent and identically distributed, and in fact are strongly correlated as discussed in Section \ref{sec_aboutErrors}.  To handle the correlated measurement errors we use Gaussian Processes \cite{rasmussen2006gaussian, DFMgp}.  The Gaussian Process takes into account the unknown but finite covariance structure by introducing a covariance matrix, $\boldsymbol{C}$ whose elements represent the correlations of each sample $i$ with every other sample $j$.  We parameterize the covariance matrix elements in the following way:

\begin{equation}
	C_{ij} = \sigma^2_{i}\delta_{ij}+a^2\exp{(-\frac{(\lambda_i-\lambda_j)^2}{2l^2})} \label{eqnGPkernel}
\end{equation}

where $\delta_{ij}$ is the Kronecker delta, and $\sigma_i$ are the independent measurement uncertainties on the $i^{\mathrm{th}}$ data point. We experimented with different values for $\sigma_i$.  When repeated DSP reference sample measurements were available, we estimated the $\sigma_i$ from the dispersion around mean-subtracted, detrended transmission spectra.  In other cases we simply adopted mean values from Table \ref{tabCary5000pars} for single and double beam modes respectively.  The latter assumption of homoscedasticity is fairly safe, since the correlated errors dominate in comparison to individual measurement errors.  The parameter $l$ controls the correlation length, and $a$ controls the amplitude of the noise.  We do not know $a$ nor $l$ and they are, in general, different for each measurement.  The value of $a$ should be close to the mean drift reported in Table \ref{tabCary5000pars}.  The value of $l$ could be small (tens of nm) to capture local variation attributable to atmospheric air absorption, or large (hundreds or thousands of nm) to capture global features attributable to lamp drift.  We allow both $a$ and $s$ to be free parameters in our fitting procedure.

@@

\begin{equation}
	\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma)} \propto \ln{p(d,f,a,s)} -\frac{1}{2}\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r} -\frac{1}{2}\;\ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
\end{equation}

where the residual vector $\boldsymbol{r}$ is the data $T_{obs}$ minus the model $T_{mix}$.  We produce posterior samples with Markov Chain Monte Carlo (MCMC).  Specifically, we implemented \texttt{emcee}\footnote{\url{https://github.com/dfm/emcee}}\cite{emcee}, with 32 walkers, hundreds of burn-in samples, and 600-1000 iterations.  We initialized the walkers in the vicinity of our best guess for parameters based on visual inspection of the observed normalized Transmission spectra.  Figures \ref{figVG03full}, \ref{figVG03part}, and \ref{figVG12_corner} show examples of posterior samples including corner plots of the interesting physical parameters $d$ and $f$.

@@

Figure \ref{figVG03full} shows a parameter determination for the measurement of a bond over a deep ($4000 \pm 200\;$nm) 100\% fill factor gap.  The analysis yields a best fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physicsal measurement.  Figure \ref{figVG03part} shows a fit for a region of the same bonded pair where a nominally $4000\;$nm deep, 2 mm wide trench crosses the field field scanned by the Cary 5000.  Analysis of this transmission curve gave a best fit depth of $4094 \pm 4\;$nm and a pattern fill factor of $4.6 \pm 0.1\;\%$, both consistent with the part geometry.  These two results show that the formalism performs as expected on resonant gaps several wavelengths deep.

>>>>

\begin{abstract}
Silicon direct bonding offers flexibility in the design and development of Si optics by allowing manufacturers to combine subcomponents with a potentially lossless and mechanically stable interface. The bonding process presents challenges in meeting the requirements for optical performance because air gaps at the Si interface cause large Fresnel reflections. Even small (35 nm) gaps reduce transmission through a direct bonded Si compound optic by 4\% at $\lambda = 1.25 \; \mu$m at normal incidence. Understanding the optical effects of such gaps and evaluation of various methods for preventing or eliminating them demands that we have a method for measuring not only the two-dimensional extent of the gaps but also their axial extent.  Existing methods for studying gaps provide only poor measures of this axial extent.  We describe a bond inspection method that makes use of precision transmission spectroscopy.  Using spectroscopy, we can detect and measure gaps as small as 12 nm by measuring transmission as a function of wavelength. Our approach involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes. We infer the gap axial extent and fill factor with Monte Carlo Markov Chain (MCMC) and a flexible Gaussian process noise model.  We describe the measurement and analysis techniques and demonstrate the validity of the approach by measuring bond gaps of known depths produced by microlithography.
\end{abstract}

@@

\subsection{The Si$-$Si interface gap is a low finesse Fabry-P\`{e}rot etalon}

@@

which has an average value of about 53\%.  Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.  Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqnAbsDSPtrans} in the wavelength range $1200-2500\;$nm.  If the Si wafer thickness is less than the coherence length for the given spectral bandwidth, then our assumptions break down and the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.  We do not treat this case here since our focus is on thick substrates with small gaps between the interfaces.

\begin{figure}[!htbp]
\includegraphics[width=0.40\columnwidth]{figs/etalon_trans.pdf}
\caption{\label{figAbsoluteTrans}The family of normal-incidence curves for transmission as a function of wavelength for various gap sizes in bonded Si optics.  \emph{Top panel}: Absolute transmission through a Si substrate with no gap (top denim-blue solid curve), and small gaps of axial extent 20 nm (green dashed curve), 35 nm (red dash-dotted curve), and 60 nm (purple dotted curve).  The absolute transmission through Si depends on wavelength since the refractive index depends minutely on wavelength.  The curves assume that multiple reflections within the Si substrates add incoherently as discussed in detail in Appendix \ref{sec:Append-IMRTMM}.  \emph{Middle panel}: The ``Gap Transmission'', defined as the Absolute transmission through a sample normalized by the absolute transmission through a gapless sample. The gap sizes and colors are the same as in the top panel.  \emph{Bottom panel}: ``Gap Transmission'' for larger gaps, spanning 0-3500 nm.  The 3500 nm sized gap (cyan solid line), demonstrates characteristic Fabry-P\`erot etalon fringes.}
\end{figure}

@@

The Cary 5000 has four operation modes: single front, single back, double, and double reverse.  We experimented with single front and double modes.  Double mode takes a reference spectrum simultaneously with the target spectrum.  In both single and double mode, a baseline scan is taken with no sample in the measurement holder.  We automatically divide sample spectra by the baseline spectrum.  In single beam mode, the baseline measurements must be repeated about every 10 minutes for post-process division.  In double beam mode the front beam is automatically divided by the rear beam, which corrects out lamp luminosity drift, for example.

@@

The spectral sampling ranged from 2.0 to 10.0 nm.  We used a spectral resolution of 5.0 nm.  We expect smoothly varying spectral features for all conceivable interface gaps in Si bonded wafers.  Specifically, Fabry-P\`erot fringes attributable to interfacial gaps in bonded Si will be well-sampled with 5 nm spectral resolution, so long as the interface gap is less than $\sim$100 $\mu$m.  The axial extents of such large gaps would be easily measurable by other means, like IR imaging.

@@

We also used IR imaging to inspect VG09-12 and VG15-13.  We detected IR bubbles in all 3 of these bonded Si samples.  The observable bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.  The individual IR bubble areas were as small as a few square millimeters up to $400\;\mathrm{mm}^2$.  The IR bubbles had varied morphologies.  The structure within the IR bubbles was sometimes indicative of Newton's rings, exposing the presence of gaps larger than about $\lambda/2$ up to $\sim 15 \lambda$.  As expected, the unbonded DSP wafers show no such IR bubbles.  No attempt at quantitative measurement was made on the IR images due to evidence of a non-linear detector response.  The bulk morphologies were clear regardless of the detector response.  When IR imaging was available, it was leveraged to target IR specectroscopy in locations demonstrating presence or absense of IR bubbles, as desired.  In other cases when IR imaging was not available or ignored, spectra were taken at random positions on the wafer.

@@

where $f$ is the areal fill factor of the region exhibiting a finite gap size, and $1-f$ is therefore the areal fill factor of the perfectly bonded area.  $T_e$ is defined in equation \ref{eqn:Tetalon}.  The only free parameters in Equation \ref{eqnMix} are $d$ and $f$.  Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tbl_meshPatterns}.  These synthetic imbedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secKnownGaps}).

Equipped with an observed spectrum and our mixture model, we constructed the likelihood function \cite{2013sdmm.book.....I}.  The measurement errors are strongly correlated as discussed in Section \ref{sec_aboutErrors}.  To handle the correlated measurement errors we use Gaussian process regression \cite{rasmussen2006gaussian, DFMgp}.  The Gaussian process takes into account the unknown but finite covariance structure by introducing a covariance matrix, $\boldsymbol{C}$ whose elements represent the correlations of each sample $i$ with every other sample $j$.  We parameterize the covariance matrix elements in the following way:

\begin{equation}
	C_{ij} = \sigma^2_{i}\delta_{ij}+a^2\exp{(-\frac{(\lambda_i-\lambda_j)^2}{2s^2})} \label{eqnGPkernel}
\end{equation}

where $\delta_{ij}$ is the Kronecker delta, and $\sigma_i$ are the independent measurement uncertainties on the $i^{\mathrm{th}}$ data point. We experimented with different values for $\sigma_i$.  When repeated DSP reference sample measurements were available, we estimated the $\sigma_i$ from the dispersion around mean-subtracted, detrended transmission spectra.  The parameter $s$ controls the correlation length, and $a$ controls the amplitude of the noise.  We do not know $a$ nor $s$ and they are, in general, different for each measurement.  The value of $a$ should be close to the mean drift reported in Table \ref{tabCary5000pars}.  The value of $s$ could be small (tens of nm) to capture local variation attributable to atmospheric air absorption, or large (hundreds or thousands of nm) to capture global features attributable to lamp drift.  We allow both $a$ and $s$ to be free parameters in our fitting procedure.

@@

\begin{equation}
	\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma_i)} \propto \ln{p(d,f,a,s)} -\frac{1}{2}\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r} -\frac{1}{2}\;\ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
\end{equation}

where the residual vector $\boldsymbol{r}$ is the data $T_{obs}$ minus the model $T_{mix}$.  We produce posterior samples with Markov Chain Monte Carlo (MCMC).  Specifically, we implemented \texttt{emcee}\footnote{\url{https://github.com/dfm/emcee}}\cite{emcee}, with 32 walkers, hundreds of burn-in samples, and 600-1000 iterations.  We initialized the walkers in the vicinity of our best guess for parameters based on visual inspection of the spectra.  Figures \ref{figVG03full}, \ref{figVG03part}, and \ref{figVG12_corner} show examples of posterior samples including corner plots of the interesting physical parameters $d$ and $f$.

@@

Figure \ref{figVG03full} shows a parameter determination for the measurement of a bond over a deep ($4000 \pm 200\;$nm) 100\% fill factor gap.  The analysis yields a best fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physicsal measurement.  Figure \ref{figVG03part} shows a fit for a region of the same bonded pair where a nominally $4000\;$nm deep, 2 mm wide trench crosses the field scanned by the Cary 5000.  Analysis of this transmission curve gave a best fit depth of $4094 \pm 4\;$nm and a pattern fill factor of $4.6 \pm 0.1\;\%$, both consistent with the part geometry.  These two results show that the formalism performs as expected on resonant gaps several wavelengths deep.
